---
title: "The influence of popularity on Tame Impala"
author: "Puck Voorham"
date: "2023-02-21"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      version: 4
      bootswatch: minty
---

```{r, echo=FALSE}

library(tidyverse)
library(plotly)
library(spotifyr)
library(compmus)
```

### Track level summary

```{r}
Before_Currents <-
  get_playlist_audio_features(
  "", "0zj7vBkFkrtkIy7bbiIDHT"
  ) |>
  slice(1:30) |>
  add_audio_analysis()

After_Currents <-
  get_playlist_audio_features(
  "", "2iF5St3YBr1AalBJFZK8uD"
  ) |>
  slice(1:30) |>
  add_audio_analysis()

Tame_Impala <-
  Before_Currents |>
  mutate(genre = "Before currents") |>
  bind_rows(After_Currents |> mutate(genre = "After currents"))

Tame_Impala |>
  mutate(
    timbre =
      map(
        segments,
        compmus_summarise,
        timbre,
        method = "mean"
      )
  ) |>
  select(genre, timbre) |>
  compmus_gather_timbre() |>
  ggplot(aes(x = basis, y = value, fill = genre)) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(x = "Spotify Timbre Coefficients", y = "", fill = "Genre")
``` 


*** 
In this visualization, you can see a comparison between the two groups within my chosen corpus: However, there is less variation than I expected beforehand in terms of timbre values. However, the second value (which says something about the "strength" of the high frequencies is significantly different. I am going to delve further into what exactly this difference means and whether I can recognize it myself. 

### Chordogram 
```{r}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

phf <-
  get_tidy_audio_analysis("1KcOU0q9AMXV9xi4tSzzCc") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

phf |> 
  compmus_match_pitch_template(
    chord_templates,         # Change to chord_templates if desired
    method = "manhattan",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")
```


### Does the genre of a major music phenomenon like Tame Impala change due to popularity?

Tame Impala is a famous musical project led by Kevin Parker, mostly known for its psychedelic rock. With the 2015 album Currents, it made its breakthrough to the greater public and has continued to grow in listeners over the past few years. Interestingly, best-known songs from the exploded album are quite "mild" by psychedelic-rock standards. Since this somewhat softer version of the genre seems to be doing well, I am curious to what extent a music phenomenon like Tame Impala, in genre 'adapts' after success, relative to their earlier songs. Do they make what the general public wants to hear, or do they maintain their authentic style for a smaller group of psychedelic-rock fanatics?

The groups I want to compare within this corpus will be, first, the first singles from 2008 and 2009, the first album Innerspeaker from 2010 and Lonerism from 2013. To the second group I want to name all the songs after Currents (2015). The album on which they broke through (Currents) I want to assess herein separately by genre.

The expectations I have at this moment are therefore that their earlier songs can be considered as harder, rougher, up-tempo and less "accessible" psychedelic rock, while the songs from Currents and thus everything that followed fit better within the genres alternative/indie and (synth)pop.

Songs that, in my opinion, typify both extremes in this are The Bold Arrow of Time from the album Innerspeaker (2010) as typical hard psychedelic rock versus The less I know the better from Currents (2015). Songs like Nangs and Gossip from that same album might be a bit off, as they could be better considered as ambient music and are hard to categorize in any of the aforementioned genres.

### A first visualisation

```{r}
Before_Currents <- get_playlist_audio_features("", "0zj7vBkFkrtkIy7bbiIDHT")
After_Currents <- get_playlist_audio_features("", "2iF5St3YBr1AalBJFZK8uD")

Categorical <- rbind(Before_Currents, After_Currents)

Categorical |>
  ggplot(aes(x = playlist_name, y = danceability, box = FALSE )) +
  geom_violin() + theme_minimal() +
  theme(
    plot.subtitle = element_text(size = 12, color = "white"),
    plot.background = element_rect(fill = "lightblue"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  ) 
```

*** 
In this first, relatively simple plot, the music used in this corpus is divided into two categories. The left form includes all songs by Tame Impala up to the Currents album in 2015. The right concerns all songs from 2015 (including Currents) to the present. In this plot, these categories are set out on the y-axis in "danceability". The expectation here was that earlier songs, before the popular album currents, were less danceable. This was the expectation because these songs could be considered more like Tame-Impala's somewhat inaccessible, rougher psychedelic rock. In short; music that could be perceived as less danceable than the smooth, disco- and synthpop-influenced songs from currents and other later tracks. The plot does somewhat support this expectation. It can be seen that at least the "peak" in danceability of the songs before currents is lower than the peak of the songs after currents. There is also less variation seen in the second group. First, the songs barely seem to score below a score of 0.4 in danceability, confirming the hypothesis. On the other hand, a single/some songs do seem to score higher in danceability than the songs after currents. To find out what this is due to requires further analysis. It could be that the whole variation in the first category is significantly higher because this category contains more songs.  


### A second visualisation

```{r}
scatter_plot <- ggplot(Categorical, aes(x = valence, y = energy, size = tempo, color = playlist_name)) +
  geom_point() +
  labs(
    x = "Valence",
    y = "Energy"
  ) + theme_minimal() +
  theme(
    text = element_text( color = "black"),
    plot.subtitle = element_text(size = 12, color = "black"),
    #plot.background = element_rect(fill = "lightblue"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  ) 

scatter_plot

```


***
deze plot is 


### Chromagram test

```{r}

Categorical <-
  get_tidy_audio_analysis("5JWPUEov2wlX7c0jhYZpeB") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)


Categorical |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  theme(
    text = element_text( color = "white"),
    plot.subtitle = element_text(size = 12, color = "white"),
    plot.background = element_rect(fill = "lightblue"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  ) +
  scale_fill_viridis_c()
```


*** 
Here is a chromagram of the song Lost in Yesterday. In this song from Tame Impala's latest album, there are clear characteristics of disco, synth-pop and funk in addition to psychedelic influences. 

### Chromagram test

```{r}

Categorical <-
  get_tidy_audio_analysis("79chzfFIIq7cHkqcYYORk0") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)


Categorical |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |> 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  theme(
    text = element_text( color = "white"),
    plot.subtitle = element_text(size = 12, color = "white"),
    plot.background = element_rect(fill = "lightblue"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  ) +
  scale_fill_viridis_c()
```

*** 
Does this work?


### Ceptrogram
```{r}
elephant <-
  get_tidy_audio_analysis("6qZjm61s6u8Ead9sWxCDro") |> # Change URI.
  compmus_align(beats, segments) |>                     # Change `bars`
  select(beats) |>                                      #   in all three
  unnest(beats) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

elephant |>
  compmus_gather_timbre() |>
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()
```


### Self similarity matrix: ceptogram

```{r}
phf <-
  get_tidy_audio_analysis("1KcOU0q9AMXV9xi4tSzzCc") |> # Change URI.
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
          compmus_summarise, timbre,
          method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

elephant |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")
```


### Chordogram 
```{r}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

phf <-
  get_tidy_audio_analysis("1KcOU0q9AMXV9xi4tSzzCc") |>
  compmus_align(sections, segments) |>
  select(sections) |>
  unnest(sections) |>
  mutate(
    pitches =
      map(segments,
          compmus_summarise, pitches,
          method = "mean", norm = "manhattan"
      )
  )

phf |> 
  compmus_match_pitch_template(
    chord_templates,         # Change to chord_templates if desired
    method = "manhattan",  # Try different distance metrics
    norm = "manhattan"     # Try different norms
  ) |>
  ggplot(
    aes(x = start + duration / 2, width = duration, y = name, fill = d)
  ) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(x = "Time (s)", y = "")

```



### Conclusion

***
Here I would write my overall conclusions so far






